%% Benjamin Bengfort's Resume
%% Copyright (C) 2016 Benjamin Bengfort

\documentclass[10pt,letterpaper,roman]{moderncv}

\moderncvstyle{oldstyle}
\moderncvcolor{green}

\usepackage[utf8]{inputenc}
\usepackage[scale=0.75]{geometry}

% personal data
\firstname{Benjamin}
\familyname{Bengfort}
\title{Resume}

\address{215 Alynn Way}{Queenstown, MD 21658}
\mobile{(701)~680~3095}
\email{benjamin@bengfort.com}
\homepage{bbengfort.github.io}

% Need to include LinkedIn, GitHub, etc.

\begin{document}

\makecvtitle

\section{Experience}

% Last Job, Hopefully
% CEO and Founder, Rotational Labs
\cventry{2020--present}{CEO and Founder}{Rotational Labs, LLC}{Queenstown, MD}{}{}

% Adjunct Faculty at Georgetown
\cventry{2014--present}{Faculty Director}{Georgetown University}{Washington, DC}{}{Adjunct Faculty, Data Analytics Certificate Program (CCPE). Teaching courses in the \textit{Data Science} and \textit{Advanced Data Science} certificates.}

% Vice President of Engineering, PingThings
% Data Scientist at Cortex
\cventry{2018--2020}{Chief Data Scientist}{PingThings, Inc.}{Washington, DC}{}{Deep neural modeling of sensor data for the smart grid, development of time-series databases and distributed sensor systems.}

% Data Scientist at Cortex
\cventry{2017--2018}{Data Scientist}{Cortex Building Intelligence}{Washington, DC}{}{Time-series analysis and modeling of building management sensor data for real-time recommendations to improve building energy efficiency.}


\cventry{2014--2017}{Partner}{District Data Labs}{Washington, DC}{}{Architect and develop innovative open source projects, facilitating local developer contributions and research, including Yellowbrick, visual steering for machine learning, and Baleen, a large scale corpus ingestion engine.}

% Data Scientist at Cobrain
\cventry{2013--2014}{Chief Data Scientist}{Cobrain Company}{Bethesda, MD}{}{
Developed a Global Recommendation Engine, using Collaborative Filtering algorithms as well as Active Learning and adaptive systems from a machine-learning standpoint. Graph traversal and clustering across massive data stores required distributed Graph databases: Titan, as well as strong computation in Hadoop and Python with Pandas and NumPy.}
% Entity resolution and ontological reasoning using probabilistic graphical models and RDF ontologies also factored into the work to disambiguate crawl data from a massive distributed crawler also developed at Cobrain.}

% CTO at Unbound Concepts
\cventry{2011--2013}{Chief Technology Officer}{Unbound Concepts}{Baltimore, MD}{}{Natural Language Processing across a large dataset of children’s literature using \newline{}%
Machine Learning and predictive analysis with clustering and multivariate non-linear regression\newline{}%
Big data analysis applied to the NLP and ML using Hadoop and MapReduce techniques.}%\newline{}%
% Rich web application development with Django along with JSON server for mobile apps.}

% Programmer at Tactical Network Solutions
\cventry{2010--2012}{Lead Programmer}{Tactical Network Solutions}{Columbia, MD}{}{Python software development for embedded, mobile and server applications\newline{}%
Large-scale and real-time data analysis of Petabytes of wireless packet collects.\newline{}%
% Wireless offensive computer network attack, computer network operations\newline{}%
Real time asset tracking software development with geolocation and KML toolkits}%\newline{}%
% Django web application development and Android mobile client applications development}

% CMS Analyst at OUP
\cventry{2007--2009}{CMS Analyst}{Oxford University Press}{Oxford, United Kingdom}{}{Management of custom content management solutions for electronic publishing division}%\newline{}%
% Customer representative for agile based development of OED 4.0 CD-ROM software\newline{}%
% Data and database management for XML DTDs and Schemas }

% Network Engineer, CenGen
\cventry{2003--2004}{Junior Network Engineer}{CenGen, Inc.}{Columbia, MD}{}{Network support for DARPA’s Grand Challenge \& USMC Condor}%\newline{}%
%Mobile and wireless integration, network vehicle integration}%\newline{}%
% Field support and logistics, small team leadership }

\section{Education}
\cventry{2014-2019}{PhD Computer Science}{University of Maryland, College Park}{}{}{Dissertation: Planetary Scale Data Storage (defended November 2018)}

\cventry{2008-2010}{M.S. Computer Science}{North Dakota State University}{}{}{Phi Kappa Phi, Upsilon Pi Epsilon}

\cventry{2004-2006}{B.A. Economics}{University of Maryland, College Park}{}{}{Primanum Honor Society}

\cventry{2002-2003}{English Major}{United States Naval Academy}{}{}{}

% \subsection{Master Thesis}
% \cvitem{title}{\emph{Title}}
% \cvitem{supervisors}{Supervisors}
% \cvitem{description}{Short thesis abstract}

\section{Skills}
\cvdoubleitem{Software Development}{GoLang 1.14+, gRPC, Python 3.6+, JavaScript, Java, C/C++, Kubernetes, Docker, Git}{Big Data}{Spark, Hadoop, MapReduce, HDFS, Distributed Systems, Celery, DISTIL}
\cvdoubleitem{Machine Learning}{Scikit-Learn, Yellowbrick, TensorFlow, SparkML, PyTorch, GraphX}{NLP}{NLTK, spaCy, Gensim, TextBlob, Pattern}
\cvdoubleitem{Application Development}{Go, Django, Nginx, AJAX, jQuery, Bootstrap, Flask, REST micro-services}{Databases}{PostgreSQL, BTrDB, MongoDB, BoltDB, LevelDB, Redis, Hive, Titan, Neo4j, SQLite}
% \cvdoubleitem{Networking}{Kismet, Wireshark, Mgen, 802.11, Protocol Buffers, NS2}{OS}{Mac OS X, Ubuntu Linux, Genoo, Embedded Ubuntu, Android, iOS.}

\nocite{*}
\bibliographystyle{plain}
\bibliography{selected_publications}

% \section{Volunteer Experience}
% \cvlistitem{UMD Department of Computer Science: Education Committee}
% % \cvlistitem{First Trinity Lutheran Church: Secretary of  Church Council}
% \cvlistitem{Advances in Cognitive Systems 2015: Referee and Reviewer}
% % \cvlistitem{University of Maryland Graduate Student Senator: 2014-2016}
% \cvlistitem{UMD Senate: Nominations Committee, Committee on Committees, Student Affairs Committee, Special Working Group on Senate Communications.}
% \cvlistitem{Data Community DC: Treasurer, Board Member, and Organizer for DIDC.}
% \cvlistitem{Vino Scholastico: Board member for HCC tuition fundraising event.}
% \cvlistitem{Team Fight: Ulman Cancer Foundation.}
% \cvlistitem{The Bridging Project: Math teacher for asylum seeking refugees.}
% \cvlistitem{Primannum Honor Society: organized blood drives for The Red Cross.}
% \cvlistitem{Leadership U: tutored ESL elementary aged students. }
% \cvlistitem{National Honor Society}

% \section{References}
% Available on request.

% \clearpage
% %-----       letter       ---------------------------------------------------------
% % recipient data
% \recipient{Addressee}{Company\\address line one\\address line two}
% \date{\today}
% \opening{Dear Sir or Madam,}
% \closing{Best Regards,}
% \enclosure[Attached]{curriculum vit\ae{}}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
% \makelettertitle
%
% Letter body.
%
%
%
% \makeletterclosing

\end{document}
