%Acknowledgments

\renewcommand{\baselinestretch}{2}
\small\normalsize
\hbox{\ }

\vspace{-.65in}

\begin{center}
\large{Acknowledgments}
\end{center}

\vspace{1ex}

No work of significance is accomplished purely through the efforts of a single individual. This work is of great significance to me and accordingly there are a great number of people whose influence either directly or indirectly led to its accomplishment. I recognize that everyone in my life from teachers and coaches to cousins and friends directed my path in someway to this point, but to ensure this acknowledgement is not a dissertation in and of itself, I will restrict myself to those whose credit may be more precisely attributed.

As my dedication on the prior page will attest, my family is the primary motivation for my pursuit of a graduate degree in the middle of my career, and was support that enabled me to do so. Both of my grandmothers emphasized education for their children and grandchildren as a way to elevate them from the circumstances of rural farming and immigration. Both my father and I pursued doctoral degrees and my mother, sisters, and many of my cousins have graduate degrees as a result. Starting a graduate program mid-career with children and a mortgage is not a choice to be taken lightly. My parents, my mother- and father-in-law, and my wife actively supported me in ways small and large and I can't thank them enough. To my children, Hank and Irena, thank you for calling me Dr. Papi even after many nights of going to bed without seeing me, and weekends watching TV while I worked on the couch (even if you are skeptical of my medical knowledge).

A special note of gratitude must be made to my wife, Jacquelyn Bengfort, whose unwavering support was superhuman, bordering on saintlike. She tolerated me through my long hours, absences, and absentmindedness; holding down the fort without a single complaint and with only words of encouragement. She picked up dropped balls and tied up loose ends, working as hard if not harder than I was to ensure that I could achieve one of my life goals.

When I left his class after my final before winter break, I remarked that all I wanted for Christmas was for Pete Keleher to email me asking if I'd like to be his graduate student. Although I didn't get that particular wish, I was thrilled when he agreed to advise me and we just clicked into our working relationship with energy and humor. I can't thank him enough for his insight and direction toward this dissertation, it would not have happened without him.

Pete has the distinction of being my last professor, though not the first and other teachers and professors deserve recognition as well. Jim Reggia, whose office was just down the hall, always enquired about my progress and offered words of support. Amol Deshpande and Dan Abadi helped me refine my work to a greater degree of precision. Michael Cox and Don Perlis brought me to the University of Maryland and supported my early doctoral work. Sergei Nirenburg convinced me to undertake the doctoral program and gave me one of the hardest decisions of my life -- to join him at RPI or to stay in DC at College Park. Xiaojiang Du, Max Zhang, and Kendall Nygard all supported me in my first computer science graduate degree.

To my lab mates and co-conspirators, thank you so much for your humor, attention, revisions, console jockeying, experiment running, data analysis, lunches, and rubber ducking. Allen, I'm sorry I convinced you to ruin your life with a graduate degree, but I'm glad we got the chance to learn together. Kostas, I'm sorry about your glasses, but without our catches in the parking lot I probably would have lost my mind. Rebecca, I'm sorry that you had to witness the sparrow that had to give its life to that eagle as a portent of my defense, thank you for helping me stay tough to the end.

Travel to conferences to present papers and attend workshops related to my graduate work was supported by the John Gannon fellowship funds, ICDCS travel committee, the Ohio State University, and Georgia Tech University. Special computing equipment and cloud environments for wide-area experiments were supported the Department of Computer Science, University of Maryland.

Thank you to my defense and proposal committee members, Dave Levin, Amol Deshpande, Daniel Abadi, Derek Richardson, and Jim Reggia for your thorough and insightful review of my work. Thank you also to Tom Hurst, Jenny Story, Regis Boykin, Stephanie Peters, Sharon McElroy and Brenda Chick without whom I would probably be lost somewhere in the bowels of A.V. Williams tied up in red tape and without heat or a desk.
